---
title: "Clase 23. Modelos de Cambio de Regímen (TAR)"
author: "Benjamin Oliva & Omar Alfaro-Rivera"
date: "30/11/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

### Dependencies and Setup
```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages('astsa')
#install.packages("tsDyn") #if not yet installed
#
library(astsa)
library(tsDyn)

```

# Data:
Monthly rates of deaths due to flu in the United States for 11 years

```{r echo = FALSE}

flu <- read.delim("flu.txt")

head(flu)

```

# Convert time series

```{r}

flu <- ts(flu)

D_flu = diff(flu, lag = 1)

```

# Plotting

```{r}

plot(flu, type = "b", col = "darkred", ylab = "", 
     main = "Monthly rates of deaths due to flu in the United States")

plot(D_flu, type="b", col = "darkred", ylab = "", 
     main = "Diff Monthly rates of deaths due to flu in the United States")

```

# Primer aproximación:

```{r}

table = ts.intersect(D_flu, 
                     lag1 = lag(D_flu, -1), 
                     lag2 = lag(D_flu, -2), 
                     lag3 = lag(D_flu, -3), 
                     lag4 = lag(D_flu, -4))

table

```

##Regression for values below the threshold

```{r}

x = table[, 1]
P = table[, 2:5]
c = .05 ## Threshold value

less = (P[,1] < c)

x1 = x[less]

P1 = P[less,]

```

```{r}
x1
```
```{r}
P1

```

```{r}

out1 = lm(x1 ~ P1[,1] + P1[,2] + P1[,3] + P1[,4])

summary(out1)

```

## Regression for values above the threshold

```{r}

greater = (P[,1]>=c)

x2 = x[greater]

P2 = P[greater,]

```

```{r}

out2 = lm(x2 ~ P2[,1] + P2[,2] + P2[,3] + P2[,4])

summary(out2)

```

## Residuals

```{r}

res1 = residuals(out1)
res2 = residuals(out2)

less[less==1] = res1
greater[greater==1] = res2

resid = less + greater
acf2(resid)

```

## Predicted values

```{r}

less = (P[,1] < c)

greater = (P[,1] >= c)

fit1 = predict(out1)

fit2 = predict(out2)

less[less==1] = fit1

greater[greater==1] = fit2

fit = less + greater

plot(D_flu, type="b", col = "darkred", ylab = "", 
     main = "Diff Monthly rates of deaths due to flu in the United States")
lines(fit, col = "darkblue", lty="dashed")

```

# The tsDyn package in R has simplified this code into a handful of steps:

```{r}

#?setar

D_flu_tar4_05 <- setar(D_flu, m = 4, thDelay = 0, th = 0.05) 

summary(D_flu_tar4_05) 

```

```{r}

plot(D_flu_tar4_05)

```

### If we do not provide a threshold to the th option, setar searches over a grid to choose a threshold ~ 0.038:

```{r}

D_flu_tar4 <- setar(D_flu, m = 4, thDelay = 0)

summary(D_flu_tar4)

plot(D_flu_tar4)

```
