---
title: "Clase 21. Datos Panel -- Raíces unitarias para panel"
author: "Benjamin Oliva & Omar Alfaro-Rivera"
date: "30/11/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

### Dependencies and Setup
```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("plm")
#
library(plm)

```

# Download data from library PLM

```{r echo = FALSE}

data("EmplUK", package="plm")

data("Produc", package="plm")

data("Grunfeld", package="plm")

data("Wages", package="plm")

```

# Describe data

Grunfeld data (Grunfeld 1958) comprising 20 annual observations on the three variables real gross investment (invest), real value of the firm (value), and real value of the capital stock (capital) for 10 large US firms for the years 1935–1954

```{r}

names(Grunfeld)

```

```{r}

head(Grunfeld)

```

```{r}

Invest <- data.frame(split( Grunfeld$inv, Grunfeld$firm )) # individuals in columns

names(Invest)

```

```{r}

names(Invest) <- c("Firm_1", "Firm_2", "Firm_3", "Firm_4", "Firm_5", "Firm_6", "Firm_7",
                   "Firm_8", "Firm_9", "Firm_10")

names(Invest)

```

# Plot:

```{r}

plot(Invest$Firm_1, type = "l", col = 1, ylim = c(0, 1500), lty = 1,
     xlab = "Tiempo", ylab = "Real gross investment")
lines(Invest$Firm_2, type = "l", col = 2, lty = 2)
lines(Invest$Firm_3, type = "l", col = 3, lty = 1)
lines(Invest$Firm_4, type = "l", col = 4, lty = 2)
lines(Invest$Firm_5, type = "l", col = 5, lty = 1)
lines(Invest$Firm_6, type = "l", col = 6, lty = 2)
lines(Invest$Firm_7, type = "l", col = 7, lty = 1)
lines(Invest$Firm_8, type = "l", col = 8, lty = 2)
lines(Invest$Firm_9, type = "l", col = 9, lty = 1)
lines(Invest$Firm_10, type = "l", col = 10, lty = 2)
legend("topleft", legend=c("Firm_1", "Firm_2", "Firm_3", "Firm_4", "Firm_5",
                           "Firm_6", "Firm_7", "Firm_8", "Firm_9", "Firm_10"),
       col = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), lty = 1:2)

```

# Unit Root Test:

Test specifies the type of test to be performed among Levin et al. (2002), Im et al. (2003), Maddala and Wu (1999) and Hadri (2000)

## Consider Levin et al. (2002)

```{r}

purtest(log(Invest), test = "levinlin", exo = "intercept", 
        lags = "AIC", pmax = 4)

```

## Same via:

```{r}

ts_LInvest <- ts(log(Invest), start = 1935, end = 1954, freq = 1)

ts_DLInvest <- diff(ts(log(Invest), start = 1935, end = 1954, freq = 1), 
                     lag = 1, differences = 1)

```

```{r}

purtest(ts_LInvest, test = "levinlin", exo = "intercept", 
        lags = "AIC", pmax = 4)

summary(purtest(ts_LInvest, test = "levinlin", exo = "intercept", 
        lags = "AIC", pmax = 4))

```

```{r}

purtest(ts_DLInvest, test = "levinlin", exo = "intercept", 
        lags = "AIC", pmax = 4)

summary(purtest(ts_DLInvest, test = "levinlin", exo = "intercept", 
        lags = "AIC", pmax = 4))

```

# Consider Im-Pesaran-Shin Unit-Root Test (2003)

```{r}

purtest(ts_LInvest, test = "ips", exo = "intercept", 
        lags = "AIC", pmax = 4)

summary(purtest(ts_LInvest, test = "ips", exo = "intercept", 
                lags = "AIC", pmax = 4))

```

```{r}
purtest(ts_DLInvest, test = "ips", exo = "intercept", 
        lags = "AIC", pmax = 4)

summary(purtest(ts_DLInvest, test = "ips", exo = "intercept", 
                lags = "AIC", pmax = 4))

```
