---
title: "Modelo Dif-in-Dif para el INPC de análsis clínicos (update nov-2022)"
author: "Benjamín Oliva"
date: "23/11/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

### Dependencies and Setup
```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages(stats)
#install.packages("lmtest")
#install.packages("foreign")
#install.packages("readxl")
#install.packages("plm")
#install.packages("tidyr")
# 
library(ggplot2)
library(dplyr)
library(stats)
library(lmtest)
library(foreign)
library(readxl)
library(plm)
library(tidyr)

options(scipen = 4)

```

# Importamos Datos desde un archivo de Excel:

Los datos están en formato panel incluyen:

* INPC (Indice Nacional de Precios al Consumidor (2QJul2018 = 100)) para un tipo de servicio
Source: https://www.inegi.org.mx/programas/inpc/2018/#Tabulados

* Dummy variable to indicate the time when the treatment started

* Dummy variable to identify the group exposed to the treatment

* Pend: Create an interaction between time and treated and add trend variable

* OTRAS variables de control

* $y = x \beta_0 + \varepsilon$

```{r echo = FALSE}

Data = read_excel("baseModeloINPC_Oct2022.xlsx", sheet = "Dummys")

dim(Data)

names(Data) <- c( "Fecha", "Mes",  "Anio", "Ciudad", "INPC", "Dummy_FIRMA", "Marca_01", "Marca_02", "Marca_03",
                  "Marca_04", "Marca_05", "Marca_06", "Marca_07", "Marca_08", "Marca_09", "Marca_10",	"Marca_11")

head(Data)

```

# Selección de ciudades:
```{r}

Data <- Data[ which( Data$Ciudad != 'Atlacomulco, Edo. de Méx.' & 
                     Data$Ciudad != 'Cancún, Q.R.' &
                     Data$Ciudad != 'Coatzacoalcos, Ver.' &
                     Data$Ciudad != 'Esperanza, Son.' &
                     Data$Ciudad != 'Izúcar de Matamoros, Pue.' &
                     Data$Ciudad != 'Pachuca, Hgo.' &
                     Data$Ciudad != 'Tuxtla Gutiérrez, Chis.' &
                     Data$Ciudad != 'Zacatecas, Zac.' ), ]

```

# Tranformaciones:

```{r}
# Log de INPC
Data$LINPC <- log( as.numeric(Data$INPC) , base = exp(1))

# Volvemos factor a la fecha
Data$Periodo <- factor(Data$Fecha, order = TRUE,
                       levels = c( "Jul 2002", "Ago 2002", "Sep 2002", "Oct 2002", "Nov 2002", "Dic 2002",
                                   "Ene 2003", "Feb 2003", "Mar 2003", "Abr 2003", "May 2003", "Jun 2003",
                                   "Jul 2003", "Ago 2003", "Sep 2003", "Oct 2003", "Nov 2003", "Dic 2003",
                                   "Ene 2004", "Feb 2004", "Mar 2004", "Abr 2004", "May 2004", "Jun 2004",
                                   "Jul 2004", "Ago 2004", "Sep 2004", "Oct 2004", "Nov 2004", "Dic 2004",
                                   "Ene 2005", "Feb 2005", "Mar 2005", "Abr 2005", "May 2005", "Jun 2005",
                                   "Jul 2005", "Ago 2005", "Sep 2005", "Oct 2005", "Nov 2005", "Dic 2005",
                                   "Ene 2006", "Feb 2006", "Mar 2006", "Abr 2006", "May 2006", "Jun 2006",
                                   "Jul 2006", "Ago 2006", "Sep 2006", "Oct 2006", "Nov 2006", "Dic 2006",
                                   "Ene 2007", "Feb 2007", "Mar 2007", "Abr 2007", "May 2007", "Jun 2007",
                                   "Jul 2007", "Ago 2007", "Sep 2007", "Oct 2007", "Nov 2007", "Dic 2007",
                                   "Ene 2008", "Feb 2008", "Mar 2008", "Abr 2008", "May 2008", "Jun 2008",
                                   "Jul 2008", "Ago 2008", "Sep 2008", "Oct 2008", "Nov 2008", "Dic 2008",
                                   "Ene 2009", "Feb 2009", "Mar 2009", "Abr 2009", "May 2009", "Jun 2009",
                                   "Jul 2009", "Ago 2009", "Sep 2009", "Oct 2009", "Nov 2009", "Dic 2009",
                                   "Ene 2010", "Feb 2010", "Mar 2010", "Abr 2010", "May 2010", "Jun 2010",
                                   "Jul 2010", "Ago 2010", "Sep 2010", "Oct 2010", "Nov 2010", "Dic 2010",
                                   "Ene 2011", "Feb 2011", "Mar 2011", "Abr 2011", "May 2011", "Jun 2011",
                                   "Jul 2011", "Ago 2011", "Sep 2011", "Oct 2011", "Nov 2011", "Dic 2011",
                                   "Ene 2012", "Feb 2012", "Mar 2012", "Abr 2012", "May 2012", "Jun 2012",
                                   "Jul 2012", "Ago 2012", "Sep 2012", "Oct 2012", "Nov 2012", "Dic 2012",
                                   "Ene 2013", "Feb 2013", "Mar 2013", "Abr 2013", "May 2013", "Jun 2013",
                                   "Jul 2013", "Ago 2013", "Sep 2013", "Oct 2013", "Nov 2013", "Dic 2013",
                                   "Ene 2014", "Feb 2014", "Mar 2014", "Abr 2014", "May 2014", "Jun 2014",
                                   "Jul 2014", "Ago 2014", "Sep 2014", "Oct 2014", "Nov 2014", "Dic 2014",
                                   "Ene 2015", "Feb 2015", "Mar 2015", "Abr 2015", "May 2015", "Jun 2015",
                                   "Jul 2015", "Ago 2015", "Sep 2015", "Oct 2015", "Nov 2015", "Dic 2015",
                                   "Ene 2016", "Feb 2016", "Mar 2016", "Abr 2016", "May 2016", "Jun 2016",
                                   "Jul 2016", "Ago 2016", "Sep 2016", "Oct 2016", "Nov 2016", "Dic 2016",
                                   "Ene 2017", "Feb 2017", "Mar 2017", "Abr 2017", "May 2017", "Jun 2017",
                                   "Jul 2017", "Ago 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dic 2017",
                                   "Ene 2018", "Feb 2018", "Mar 2018", "Abr 2018", "May 2018", "Jun 2018",
                                   "Jul 2018", "Ago 2018", "Sep 2018", "Oct 2018", "Nov 2018", "Dic 2018",
                                   "Ene 2019", "Feb 2019", "Mar 2019", "Abr 2019", "May 2019", "Jun 2019",
                                   "Jul 2019", "Ago 2019", "Sep 2019", "Oct 2019", "Nov 2019", "Dic 2019",
                                   "Ene 2020", "Feb 2020", "Mar 2020", "Abr 2020", "May 2020", "Jun 2020",
                                   "Jul 2020", "Ago 2020", "Sep 2020", "Oct 2020", "Nov 2020", "Dic 2020",
                                   "Ene 2021", "Feb 2021", "Mar 2021", "Abr 2021", "May 2021", "Jun 2021",
                                   "Jul 2021", "Ago 2021", "Sep 2021", "Oct 2021", "Nov 2021", "Dic 2021",
                                   "Ene 2022", "Feb 2022", "Mar 2022", "Abr 2022", "May 2022", "Jun 2022",
                                   "Jul 2022", "Ago 2022", "Sep 2022", "Oct 2022", "Nov 2022", "Dic 2022" ))

```

# Plot

```{r echo = TRUE}

Data %>%
  ggplot( aes(x = Periodo, y = LINPC, group = Ciudad, color = Ciudad )) +
        geom_line() +
        theme(legend.title = element_text(color = "black", size = 8),
          legend.text = element_text(color = "black", size = 5 )) +
        theme(legend.position="bottom") +
        theme(axis.text.x = element_text( size = 8, angle = 90))


```

# Pruebas de Raíces Unitarias

## Change to format wide

```{r echo = FALSE}

LINPC <- spread(Data[c(1,4,18)], key = Ciudad, value = LINPC)

head(LINPC)

```

# Time Series:

```{r}

ts_LINPC <- ts( LINPC[c( 2:48 )], 
               start = c(2002, 7),
               freq = 12)

ts_DLINPC <- diff(ts( LINPC[c( 2:48 )], 
                     start = c(2002, 7),
                     freq = 12), 
                  lag = 1, 
                  differences = 1)

```
## Unit Root Test:

Test specifies the type of test to be performed among Levin et al. (2002), Im et al. (2003), Maddala and Wu (1999) and Hadri (2000)

Optimal lags:
p = int{4*(T/100)^(1/4)}
   = int{4*(244/100)^(1/4)}
   = int{4.99}
   = 4

## Consider Levin et al. (2002)

```{r}

purtest(ts_LINPC, test = "levinlin", exo = "trend", # exo = c("none", "intercept", "trend"),
        lags = "AIC", pmax = 4)

```

```{r}

purtest(ts_DLINPC, test = "levinlin", exo = "trend", # exo = c("none", "intercept", "trend"),
        lags = "AIC", pmax = 4)

```

## Consider Choi (2001)

```{r}

purtest(ts_LINPC, test = "logit", exo = "trend", # exo = c("none", "intercept", "trend"),
        lags = "AIC", pmax = 4)

```

```{r}

purtest(ts_DLINPC, test = "logit", exo = "trend", # exo = c("none", "intercept", "trend"),
        lags = "AIC", pmax = 4)

```

# Consider Hadri (2000)

```{r}

purtest(ts_LINPC, test = "hadri", exo = "trend", # exo = c("none", "intercept", "trend"),
        lags = "AIC", pmax = 4)

```

```{r}

purtest(ts_DLINPC, test = "hadri", exo = "trend", # exo = c("none", "intercept", "trend"),
        lags = "AIC", pmax = 4)


```

# Estimación PANEL con TREND

```{r}
# Modelos a REPORTAR
fixed <- plm( LINPC ~ Dummy_FIRMA + Marca_01 + Marca_02 + Marca_03 + Marca_04 + Marca_05 + 
              Marca_06 + Marca_07 + Marca_08 + Marca_09 + Marca_10 + Marca_11 + as.numeric(Periodo),
             data = Data, 
             index=c("Ciudad", "Periodo"), 
             model = "within")

summary(fixed)

```

```{r}
# Modelos a REPORTAR
fixed <- plm( LINPC ~ Dummy_FIRMA + as.numeric(Periodo),
             data = Data, 
             index=c("Ciudad", "Periodo"), 
             model = "within")

summary(fixed)

```

