---
title: "Clase 20. Modelo ADRL"
author: "Benjamin Oliva & Omar Alfaro-Rivera"
date: "30/11/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

### Dependencies and Setup
```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("zoo")
#install.packages("xts")
#install.packages("ARDL")
# 
library(zoo) 
library(xts) 
library(ARDL)

```

# DESCRIPCIÓN DEL PROBLEMA

Supongamos que queremos modelar el logaritmo de dinero (M2) como una función de LRY (logarithm of real income), IBO (bond rate) e IDE (bank deposit rate). 

* El problema es que la aplicación de una regresión de MCO en datos no estacionarios daría lugar a una regresión espúria. 

* Los parámetros estimados serían consistentes solo si las series estuvieran cointegradas.
 

# Importamos Datos desde un dataset de R:

A data frame with 55 rows and 5 variables. Time period from 1974:Q1 until 1987:Q3.

LRM: logarithm of real money, M2

LRY: logarithm of real income

LPY: logarithm of price deflator

IBO: bond rate

IDE: bank deposit rate

```{r echo = FALSE}

data(denmark)
#?denmark # see for more information
names(denmark)

```

# Procedimiento:

## 1. Calculamos un auto ADRL para determinar la combinación óptima de rezagos.

```{r}

models <- auto_ardl(LRM ~ LRY + IBO + IDE, data = denmark, max_order = 5)

names(models)

```

```{r}

models$top_orders

```

```{r}

models$best_order

```

```{r}
models$best_model

```

```{r}
BestMod <- models$best_model

summary(BestMod)

```

## 2. UECM (Unrestricted Error Correction Model) of the underlying ARDL.
```{r}

UECM_BestMod <- uecm(BestMod)

summary(UECM_BestMod)

```

## 3. RECM (Restricted Error Correction Model) of the underlying ARDL
Obs: allowing the constant to join the short-run relationship (case 2), instead of the long-run (case 3)

```{r}

RECM_BestMod <- recm(UECM_BestMod, case = 2)

summary(RECM_BestMod)

```

## 4. long-run levels relationship (cointegration) 

```{r}

bounds_f_test(BestMod, case = 2)

```

## 5. Long-run multipliers (with standard errors, t-statistics and p-values)

```{r}

multipliers(BestMod)

```

```{r}

Result <- coint_eq(BestMod, case = 2)

```

# Make the plot

```{r}

Datos <- cbind.zoo(LRM = denmark[,"LRM"], Result)

Datos <- xts(Datos)

plot(Datos, legend.loc = "right")

```
#